---
title: "Donatelli et.al. 2021: Stats and Figures"
output: html_notebook
---

      You need Rtools I think to install some of these packages
https://cran.r-project.org/bin/windows/Rtools/rtools43/rtools.html

      You also need BiocManager to install mixOmics
https://cran.r-project.org/web/packages/BiocManager/vignettes/BiocManager.html

Libraries
```{r message=FALSE, warning=FALSE, include=FALSE}

# stats things
library(caret)
## BiocManager::install('mixOmics')
library(mixOmics)
## BiocManager::install("ropls")
library(ropls)
library(bbmle)   #for AICtab()
library(car)     #for Anova()
library(lme4)

# organizing
library(Rmisc)
library(readxl)
library(tidyverse)
library(here)
library(broom)
library(tidyr)
library(dplyr)

# for plotting
library(ggplot2)
library(ggthemes)
library(RColorBrewer)
library(cowplot)

# for plotting on a tree
library("ape")
library("phytools")
```

# Functions
```{r}
LDAModel <- function(data){
  names(data)[1] <- "Mode"
  data<-na.omit(data)
  as.factor(data$Mode)
  
  set.seed(123)
  training.samples <- data$Mode %>%
  createDataPartition(p = 0.8, list = FALSE)
  train.data <- data[training.samples, ]
  test.data <- data[-training.samples, ]
  
  # Estimate preprocessing parameters
  preproc.param <- train.data %>% 
  preProcess(method = c("center", "scale"))

  # Transform the data using the estimated parameters
  train.transformed <- preproc.param %>% predict(train.data)
  test.transformed <- preproc.param %>% predict(test.data)

  # Fit the model
  model <- lda(Mode~., data = train.transformed)
  # Make predictions
  predictions <- predict(model, test.transformed, method = "predictive")
  # Model accuracy
  modelAccuracy <- mean(predictions$class==test.transformed$Mode)
  prop = round((model$svd^2/sum(model$svd^2))*100, digits = 2)
  
  out <- list(model, train.transformed, modelAccuracy, prop)
  return(out)
}

PLSDAModel <- function(data){
  names(data)[1] <- "Mode"
  data<-na.omit(data)
  data$Mode<- as.factor(data$Mode)
  
  training.samples <- data$Mode %>%
    createDataPartition(p = 0.85, list = FALSE)
    train.data <- data[training.samples, ]
    test.data <- data[-training.samples, ]
  
  # opls.Data <- opls(as.matrix(train.data[2:ncol(train.data)]), train.data$Mode)
  plsda.data<- plsda(as.matrix(train.data[2:ncol(train.data)]), train.data$Mode)
  
  predictions <- predict(plsda.data, test.data[2:ncol(test.data)], method = "predictive")
  
  modelAccuracy <- mean(predictions$class$mahalanobis.dist[,1]==test.data$Mode)
  
  out<-list(plsda.data, predictions, modelAccuracy)
  
}
```

# Read In Data
```{r message=FALSE, warning=FALSE, include=FALSE}
# Material testing data for real fish
vertData <- read_csv("LayanneFHL2020.csv")
vertData[vertData == 0] <- NA

# Measuremtnts from CT scans
vertMeasurementsAll <- read_csv("MergedWithNewCoordinate.csv")
vertMeasurementsAll[vertMeasurementsAll == 0] <- NA

my.data <- read_excel("MasterVert_Measurements_forCassandra.xlsm", 
    sheet = "VertMeasurements")
my.data <- my.data[c(1:4, 11:16,33:40)]
my.data <- na.omit(subset(my.data, Indiv == 1))
```

# Models for bending real verts
All Bending Data
            Sum Sq  Df F value    Pr(>F)    
PercSquish  1.2145   1  17.427 4.795e-05 ***
Residuals  11.6382 167 

Real Verts
            Sum Sq Df F value    Pr(>F)    
PercSquish 3.00728  1  118.43 3.091e-13 ***
Residuals  0.96496 38  
```{r}
bendingModelAll <- lm(Npermm ~ PercSquish, data = vertData)
  # Anova(bendingModelAll)

realVerts <- filter(vertData, Type == "Real")
bendingModelReal <- lm(Npermm ~ PercSquish, data = realVerts)
  # Anova(bendingModelReal)

vertMeasurementsAll = mutate(vertMeasurementsAll, predNmm = coef(bendingModelReal)[1] + coef(bendingModelReal)[2]*PercSquish)
my.data = mutate(my.data, predNmm = coef(bendingModelReal)[1] + coef(bendingModelReal)[2]*PercSquish)
```


# Subset dataset for LDA to avoid colinearity
```{r}
  vertMeasurements <- filter(vertMeasurementsAll, Position != "20")
  vertMeasurements <- filter(vertMeasurements, Position != "30")
  vertMeasurements <- filter(vertMeasurements, Position != "40")
  vertMeasurements <- filter(vertMeasurements, Position != "60")
  vertMeasurements <- filter(vertMeasurements, Position != "80")
```

# Calculate coefifients of fit lines along the body
```{r}
library(dplyr)

VertAmp <- filter(vertMeasurements, Amp != "NA")
VertWob <- filter(vertMeasurements, Wobble != "NA")

 PLSdataFrameAll <- vertMeasurementsAll %>%
    group_by(Species, Individual) %>% #for each individual...
    summarize(X.CBL.I = lm(CBL ~ poly(Position, 2)) %>% coef() %>% .[1],  #fit a polynomial model, then get the coeficient in slot 1
              X.CBL.p1 = lm(CBL ~ poly(Position, 2)) %>% coef() %>% .[2], #...slot 2
              X.CBL.p2 = lm(CBL ~ poly(Position, 2)) %>% coef() %>% .[3], #...slot 3
              X.Dant.I = lm(Dant ~ poly(Position, 2)) %>% coef() %>% .[1],
              X.Dant.p1 = lm(Dant ~ poly(Position, 2)) %>% coef() %>% .[2],
              X.Dant.p2 = lm(Dant ~ poly(Position, 2)) %>% coef() %>% .[3],
              X.Dpost.I = lm(Dpost ~ poly(Position, 2)) %>% coef() %>% .[1],
              X.Dpost.p1 = lm(Dpost ~ poly(Position, 2)) %>% coef() %>% .[2],
              X.Dpost.p2 = lm(Dpost ~ poly(Position, 2)) %>% coef() %>% .[3],
              X.d.I = lm(d ~ poly(Position, 2)) %>% coef() %>% .[1],
              X.d.p1 = lm(d ~ poly(Position, 2)) %>% coef() %>% .[2],
              X.d.p2 = lm(d ~ poly(Position, 2)) %>% coef() %>% .[3],
              X.alphaAnt.I = lm(alphaAnt ~ poly(Position, 2)) %>% coef() %>% .[1],
              X.alphaAnt.p1 = lm(alphaAnt ~ poly(Position, 2)) %>% coef() %>% .[2],
              X.alphaAnt.p2 = lm(alphaAnt ~ poly(Position, 2)) %>% coef() %>% .[3],
              X.alphaPost.I = lm(alphaPost ~ poly(Position, 2)) %>% coef() %>% .[1],
              X.alphaPost.p1 = lm(alphaPost ~ poly(Position, 2)) %>% coef() %>% .[2],
              X.alphaPost.p2 = lm(alphaPost ~ poly(Position, 2)) %>% coef() %>% .[3],
              X.perc.I = lm(PercSquish ~ poly(Position, 2)) %>% coef() %>% .[1],
              X.perc.p1 = lm(PercSquish ~ poly(Position, 2)) %>% coef() %>% .[2],
              X.perc.p2 = lm(PercSquish ~ poly(Position, 2)) %>% coef() %>% .[3],
              Mode = first(Mode),
              Habitat = first(Habitat),
              Shape = first(Shape))
 
PLSdataFrameAmp <- VertAmp %>%
    group_by(Species, Individual) %>% #for each individual...
    summarize(X.CBL.I = lm(CBL ~ poly(Position, 2)) %>% coef() %>% .[1],  #fit a polynomial model, then get the coeficient in slot 1
              X.CBL.p1 = lm(CBL ~ poly(Position, 2)) %>% coef() %>% .[2], #...slot 2
              X.CBL.p2 = lm(CBL ~ poly(Position, 2)) %>% coef() %>% .[3], #...slot 3
              X.Dant.I = lm(Dant ~ poly(Position, 2)) %>% coef() %>% .[1],
              X.Dant.p1 = lm(Dant ~ poly(Position, 2)) %>% coef() %>% .[2],
              X.Dant.p2 = lm(Dant ~ poly(Position, 2)) %>% coef() %>% .[3],
              X.Dpost.I = lm(Dpost ~ poly(Position, 2)) %>% coef() %>% .[1],
              X.Dpost.p1 = lm(Dpost ~ poly(Position, 2)) %>% coef() %>% .[2],
              X.Dpost.p2 = lm(Dpost ~ poly(Position, 2)) %>% coef() %>% .[3],
              X.d.I = lm(d ~ poly(Position, 2)) %>% coef() %>% .[1],
              X.d.p1 = lm(d ~ poly(Position, 2)) %>% coef() %>% .[2],
              X.d.p2 = lm(d ~ poly(Position, 2)) %>% coef() %>% .[3],
              X.alphaAnt.I = lm(alphaAnt ~ poly(Position, 2)) %>% coef() %>% .[1],
              X.alphaAnt.p1 = lm(alphaAnt ~ poly(Position, 2)) %>% coef() %>% .[2],
              X.alphaAnt.p2 = lm(alphaAnt ~ poly(Position, 2)) %>% coef() %>% .[3],
              X.alphaPost.I = lm(alphaPost ~ poly(Position, 2)) %>% coef() %>% .[1],
              X.alphaPost.p1 = lm(alphaPost ~ poly(Position, 2)) %>% coef() %>% .[2],
              X.alphaPost.p2 = lm(alphaPost ~ poly(Position, 2)) %>% coef() %>% .[3],
              X.perc.I = lm(PercSquish ~ poly(Position, 2)) %>% coef() %>% .[1],
              X.perc.p1 = lm(PercSquish ~ poly(Position, 2)) %>% coef() %>% .[2],
              X.perc.p2 = lm(PercSquish ~ poly(Position, 2)) %>% coef() %>% .[3],
              Mode = first(Mode),
              Habitat = first(Habitat),
              Shape = first(Shape),
              Y.Amp.I = lm(Amp ~ poly(Position, 1)) %>% coef() %>% .[1],
              Y.Amp.p1 = lm(Amp ~ poly(Position, 1)) %>% coef() %>% .[2],
              Y.Amp.mean = mean(Amp))

 PLSdataFrameWob <- VertWob %>%
    group_by(Species, Individual) %>% #for each individual...
    summarize(X.CBL.I = lm(CBL ~ poly(Position, 2)) %>% coef() %>% .[1],  #fit a polynomial model, then get the coeficient in slot 1
              X.CBL.p1 = lm(CBL ~ poly(Position, 2)) %>% coef() %>% .[2], #...slot 2
              X.CBL.p2 = lm(CBL ~ poly(Position, 2)) %>% coef() %>% .[3], #...slot 3
              X.Dant.I = lm(Dant ~ poly(Position, 2)) %>% coef() %>% .[1],
              X.Dant.p1 = lm(Dant ~ poly(Position, 2)) %>% coef() %>% .[2],
              X.Dant.p2 = lm(Dant ~ poly(Position, 2)) %>% coef() %>% .[3],
              X.Dpost.I = lm(Dpost ~ poly(Position, 2)) %>% coef() %>% .[1],
              X.Dpost.p1 = lm(Dpost ~ poly(Position, 2)) %>% coef() %>% .[2],
              X.Dpost.p2 = lm(Dpost ~ poly(Position, 2)) %>% coef() %>% .[3],
              X.d.I = lm(d ~ poly(Position, 2)) %>% coef() %>% .[1],
              X.d.p1 = lm(d ~ poly(Position, 2)) %>% coef() %>% .[2],
              X.d.p2 = lm(d ~ poly(Position, 2)) %>% coef() %>% .[3],
              X.alphaAnt.I = lm(alphaAnt ~ poly(Position, 2)) %>% coef() %>% .[1],
              X.alphaAnt.p1 = lm(alphaAnt ~ poly(Position, 2)) %>% coef() %>% .[2],
              X.alphaAnt.p2 = lm(alphaAnt ~ poly(Position, 2)) %>% coef() %>% .[3],
              X.alphaPost.I = lm(alphaPost ~ poly(Position, 2)) %>% coef() %>% .[1],
              X.alphaPost.p1 = lm(alphaPost ~ poly(Position, 2)) %>% coef() %>% .[2],
              X.alphaPost.p2 = lm(alphaPost ~ poly(Position, 2)) %>% coef() %>% .[3],
              X.perc.I = lm(PercSquish ~ poly(Position, 2)) %>% coef() %>% .[1],
              X.perc.p1 = lm(PercSquish ~ poly(Position, 2)) %>% coef() %>% .[2],
              X.perc.p2 = lm(PercSquish ~ poly(Position, 2)) %>% coef() %>% .[3],
              Mode = first(Mode),
              Habitat = first(Habitat),
              Shape = first(Shape),
              Y.Wob.I = lm(Wobble ~ poly(Position, 1)) %>% coef() %>% .[1],
              Y.Wob.p1 = lm(Wobble ~ poly(Position, 1)) %>% coef() %>% .[2],
              Y.Wob.mean = mean(Wobble),
              Y.G.I = lm(G ~ poly(Position, 1)) %>% coef() %>% .[1],
              Y.G.p1 = lm(G ~ poly(Position, 1)) %>% coef() %>% .[2],
              Y.G.mean = mean(G),
              Y.GJ.I = lm(GJ ~ poly(Position, 1)) %>% coef() %>% .[1],
              Y.GJ.p1 = lm(GJ ~ poly(Position, 1)) %>% coef() %>% .[2],
              Y.GJ.mean = mean(GJ))
```

# LME on Measurements
```{r}
VertAmp <- filter(vertMeasurements, Amp != "NA")
VertWob <- filter(vertMeasurements, Wobble != "NA")

dat <- VertAmp[14:19]
scaled.dat <- scale(dat, center = FALSE)
VertAmpModel <- cbind(VertAmp[c(1:3,10)], scaled.dat)

ampLM <- lmer(Amp ~ Dpost*alphaPost+Dant*alphaAnt+CBL+d+
                (1|Individual)+(1|Species),
              VertAmpModel, na.action = na.omit)
  Anova(ampLM)

dat <- VertWob[14:19]
scaled.dat <- scale(dat, center = FALSE)
VertWobModel <- cbind(VertWob[c(1:3,11)], scaled.dat)

wobLM <- lmer(Wobble ~ Dpost*alphaPost+Dant*alphaAnt+CBL+d+
                (1|Individual)+(1|Species),
              VertWobModel, na.action = na.omit)
  Anova(wobLM)
  
VertGJModel <- cbind(VertWob[c(1:3,13)], scaled.dat)
GJLM <- lmer(GJ ~ Dpost*alphaPost+Dant*alphaAnt+CBL+d+
               (1|Individual),
              VertGJModel, na.action = na.omit)
  Anova(GJLM)
  
VertGModel <- cbind(VertWob[c(1:3,12)], scaled.dat)
GLM1 <- lm(G ~ Dpost*alphaPost+Dant*alphaAnt+CBL+d,
              VertGModel, na.action = na.omit)
  Anova(GLM1)
  
rm(dat, scaled.dat)
```

# Rotate Vert Measurement Data
```{r}
fish <-
  vertMeasurements %>%
  select(c('Species', 'Individual', 'Habitat', 'Shape','Depth','Association', 'Mode', 'Position', 'CBL', 'd', 'alphaPost', 'alphaAnt', 'Dpost', 'Dant','ASquishAnt', 'ASquishPost', 'ABoneL', 'ABoneR', 'PercSquish', 'predNmm')) %>%
  # "complete" the Pos column by filling in values that are missing for Species
  # and individuals but present for others
  complete(Position, nesting(Species, Individual)) %>%
  arrange(Species, Individual, Position)

fish <-
  fish %>%
  group_by(Species, Individual) %>%
  fill(Habitat, Shape, Mode, Depth, Association, .direction='downup') %>%
  ungroup()

fish_full <-
  fish %>%
  group_by(Species, Individual) %>%
  mutate_at(vars(d, Dpost, Dant, alphaPost, alphaAnt, CBL, ASquishAnt, ASquishPost, ABoneL, ABoneR, PercSquish, predNmm), ~ replace(.x, is.na(.x), mean(.x, na.rm=TRUE))) %>%
  ungroup()

fishRotated <-
fish_full %>%
  pivot_wider(names_from = Position, values_from = c('CBL', 'alphaPost', 'alphaAnt', 'd', 'Dpost','Dant', 'ASquishAnt', 'ASquishPost', 'ABoneL', 'ABoneR', 'PercSquish','predNmm'))

rm(fish, fish_full)
```

# LDA on Vert Measurements
```{r}
# Linear = c(?, 8:25)          Mode = 7
# Areas = c(?, 26:37)         Shape = 4
# Perc = c(?, 38:40)          Habitat = 3
# PredNmm = c(:, 41:43)

## Habitat
LDHabitat.L         <- LDAModel(fishRotated[c(3,c(8:25))])  # Linear Only
LDHabitat.L.A       <- LDAModel(fishRotated[c(3,c(8:37))])  # Linear + Area
LDHabitat.L.Pc      <- LDAModel(fishRotated[c(3,c(8:25, 38:40))])  # Linear + Perc 
LDHabitat.L.Pd      <- LDAModel(fishRotated[c(3,c(8:25, 41:43))])  # Linear + Pred 

HabitatVals <- matrix(1:12, nrow = 4, ncol = 3) 
colnames(HabitatVals) <- c("Accuracy", "LD1", "LD2")
rownames(HabitatVals) <- c("Linear", "L+Area", "L+Perc", "L+Pred")
HabitatVals[1,] = c(LDHabitat.L[[3]], LDHabitat.L[[4]][1], LDHabitat.L[[4]][2]) 
HabitatVals[2,] = c(LDHabitat.L.A[[3]], LDHabitat.L.A[[4]][1], LDHabitat.L.A[[4]][2]) 
HabitatVals[3,] = c(LDHabitat.L.Pc[[3]], LDHabitat.L.Pc[[4]][1], LDHabitat.L.Pc[[4]][2]) 
HabitatVals[4,] = c(LDHabitat.L.Pd[[3]], LDHabitat.L.Pd[[4]][1], LDHabitat.L.Pd[[4]][2]) 
HabitatVals

## Shape
LDShape.L         <- LDAModel(fishRotated[c(4,c(8:25))])  # Linear Only
LDShape.L.A       <- LDAModel(fishRotated[c(4,c(8:37))])  # Linear + Area
LDShape.L.Pc      <- LDAModel(fishRotated[c(4,c(8:25, 38:40))])  # Linear + Perc 
LDShape.L.Pd      <- LDAModel(fishRotated[c(4,c(8:25, 41:43))])  # Linear + Pred 

ShapeVals <- matrix(1:12, nrow = 4, ncol = 3) 
colnames(ShapeVals) <- c("Accuracy", "LD1", "LD2")
rownames(ShapeVals) <- c("Linear", "L+Area", "L+Perc", "L+Pred")
ShapeVals[1,] = c(LDShape.L[[3]], LDShape.L[[4]][1], LDShape.L[[4]][2]) 
ShapeVals[2,] = c(LDShape.L.A[[3]], LDShape.L.A[[4]][1], LDShape.L.A[[4]][2]) 
ShapeVals[3,] = c(LDShape.L.Pc[[3]], LDShape.L.Pc[[4]][1], LDShape.L.Pc[[4]][2]) 
ShapeVals[4,] = c(LDShape.L.Pd[[3]], LDShape.L.Pd[[4]][1], LDShape.L.Pd[[4]][2]) 
ShapeVals

## Mode
LDMode.L         <- LDAModel(fishRotated[c(7,c(8:25))])  # Linear Only
LDMode.L.A       <- LDAModel(fishRotated[c(7,c(8:37))])  # Linear + Area
LDMode.L.Pc      <- LDAModel(fishRotated[c(7,c(8:25, 38:40))])  # Linear + Perc 
LDMode.L.Pd      <- LDAModel(fishRotated[c(7,c(8:25, 41:43))])  # Linear + Pred 

ModeVals <- matrix(1:12, nrow = 4, ncol = 3) 
colnames(ModeVals) <- c("Accuracy", "LD1", "LD2")
rownames(ModeVals) <- c("Linear", "L+Area", "L+Perc", "L+Pred")
ModeVals[1,] = c(LDMode.L[[3]], LDMode.L[[4]][1], LDMode.L[[4]][2]) 
ModeVals[2,] = c(LDMode.L.A[[3]], LDMode.L.A[[4]][1], LDMode.L.A[[4]][2]) 
ModeVals[3,] = c(LDMode.L.Pc[[3]], LDMode.L.Pc[[4]][1], LDMode.L.Pc[[4]][2]) 
ModeVals[4,] = c(LDMode.L.Pd[[3]], LDMode.L.Pd[[4]][1], LDMode.L.Pd[[4]][2]) 
ModeVals

```

# LDA Plots
NOTE: If you get "lazy-load database 'keras.rdb' is corrupt" when running this section you just need to restart your R-Session
```{r}
Habitat.lda.data <- cbind(LDHabitat.L.A[[2]], predict(LDHabitat.L.A[[1]])$x)
Shape.lda.data <- cbind(LDShape.L[[2]], predict(LDShape.L[[1]])$x)
Mode.lda.data <- cbind(LDMode.L.A[[2]], predict(LDMode.L.A[[1]])$x)

ggplot(Habitat.lda.data, aes(LD1, LD2)) +
    geom_point(aes(color = Mode)) +
    theme_classic()+
    stat_ellipse(aes(color = Mode))+
    scale_color_brewer(palette="Dark2")+
    labs(color = "Habitat",
         x = paste("LD1 (", LDHabitat.L.A[[4]][1], ")", sep=""),
         y = paste("LD2 (", LDHabitat.L.A[[4]][2], ")", sep=""))+
    theme(axis.text.x = element_text(size = 14),
      text = element_text(size=20))

ggplot(Shape.lda.data, aes(LD1, LD2)) +
    geom_point(aes(color = Mode)) +
    theme_classic()+
    stat_ellipse(aes(color = Mode))+
    scale_color_brewer(palette="Dark2")+
    labs(color = "Shape",
         x = paste("LD1 (", LDShape.L.A[[4]][1], ")", sep=""),
         y = paste("LD2 (", LDShape.L.A[[4]][2], ")", sep=""))+
    theme(axis.text.x = element_text(size = 14),
      text = element_text(size=20))

ggplot(Mode.lda.data, aes(LD1, LD2)) +
    geom_point(aes(color = Mode)) +
    theme_classic()+
    stat_ellipse(aes(color = Mode))+
    scale_color_brewer(palette="Dark2")+
    labs(color = "Mode",
         x = paste("LD1 (", LDMode.L.A[[4]][1], ")", sep=""),
         y = paste("LD2 (", LDMode.L.A[[4]][2], ")", sep=""))+
    theme(axis.text.x = element_text(size = 14),
      text = element_text(size=20))
#       ggsave("Habitat.eps", width = 18, height = 8, units = "cm")
```

# LDA on Coefficients
```{r}
# Linear = c(?, 8:25)          Mode = 24
# Areas = c(?, 26:37)         Shape = 26
# Perc = c(?, 38:40)          Habitat = 25
# PredNmm = c(:, 41:43)

## Habitat
LDHabitat.Coef.Perc <- LDAModel(PLSdataFrameAll[c(25,c(3:23))])  # Linear Only
LDMode.Coef.Perc <- LDAModel(PLSdataFrameAll[c(24,c(3:23))])  # Linear Only
LDShape.Coef.Perc <- LDAModel(PLSdataFrameAll[c(26,c(3:23))])  # Linear Only

LDHabitat.Coef <- LDAModel(PLSdataFrameAll[c(25,c(3:20))])  # Linear Only
LDMode.Coef <- LDAModel(PLSdataFrameAll[c(24,c(3:20))])  # Linear Only
LDShape.Coef <- LDAModel(PLSdataFrameAll[c(26,c(3:20))])  # Linear Only

CoefVals <- matrix(1:18, nrow = 6, ncol = 3) 
colnames(CoefVals) <- c("Accuracy", "LD1", "LD2")
rownames(CoefVals) <- c("Habitat", "Mode", "Shape","HabitatPerc", "ModePerc", "ShapePerc")

CoefVals[1,] = c(LDHabitat.Coef[[3]], LDHabitat.Coef[[4]][1], LDHabitat.Coef[[4]][2]) 
CoefVals[2,] = c(LDMode.Coef[[3]], LDMode.Coef[[4]][1], LDMode.Coef[[4]][2]) 
CoefVals[3,] = c(LDShape.Coef[[3]], LDShape.Coef[[4]][1], LDShape.Coef[[4]][2]) 
CoefVals[4,] = c(LDHabitat.Coef.Perc[[3]], LDHabitat.Coef.Perc[[4]][1], LDHabitat.Coef.Perc[[4]][2]) 
CoefVals[5,] = c(LDMode.Coef.Perc[[3]], LDMode.Coef.Perc[[4]][1], LDMode.Coef.Perc[[4]][2]) 
CoefVals[6,] = c(LDShape.Coef.Perc[[3]], LDShape.Coef.Perc[[4]][1], LDShape.Coef.Perc[[4]][2])
CoefVals

scalingTable = matrix(nrow = 18, ncol = 6)
rownames(scalingTable) <- row.names(LDMode.Coef[[1]]$scaling)
colnames(scalingTable) <- c("Habitat LD1", "Habitat LD2", "Mode LD1", "Mode LD2",
                            "Shape LD1", "Shape LD2")
scalingTable[,1:2]   = LDHabitat.Coef[[1]]$scaling[,1:2]
scalingTable[,3:4]   = LDMode.Coef[[1]]$scaling[,1:2]
scalingTable[,5:6]   = LDShape.Coef[[1]]$scaling[,1:2]
write.table(scalingTable, file = "scalingTable.csv", sep = ",")

scalingTablePerc = matrix(nrow = 21, ncol = 6)
rownames(scalingTablePerc) <- row.names(LDMode.Coef.Perc[[1]]$scaling)
colnames(scalingTablePerc) <- c("P Habitat LD1", "P Habitat LD2", 
                            "P Mode LD1", "P Mode LD2", "P Shape LD1", "P Shape LD2")
scalingTablePerc[,1:2]   = LDHabitat.Coef.Perc[[1]]$scaling[,1:2]
scalingTablePerc[,3:4]  = LDMode.Coef.Perc[[1]]$scaling[,1:2]
scalingTablePerc[,5:6] = LDShape.Coef.Perc[[1]]$scaling[,1:2]
write.table(scalingTablePerc, file = "scalingTablePerc.csv", sep = ",")
```

# LDA Coefficient Plots
```{r}
Habitat.lda.data <- cbind(LDHabitat.Coef[[2]], predict(LDHabitat.Coef.Perc[[1]])$x)
Shape.lda.data <- cbind(LDShape.Coef[[2]], predict(LDShape.Coef.Perc[[1]])$x)
Mode.lda.data <- cbind(LDMode.Coef[[2]], predict(LDMode.Coef.Perc[[1]])$x)

p1.Hab<-ggplot(Habitat.lda.data, aes(LD1, LD2)) +
    geom_point(aes(color = Mode)) +
    theme_classic()+
    stat_ellipse(aes(color = Mode), type = "t", level = 0.75)+
    scale_color_brewer(palette="Dark2")+
    labs(color = "Habitat",
         x = paste("LD1 (", LDHabitat.Coef[[4]][1], ")", sep=""),
         y = paste("LD2 (", LDHabitat.Coef[[4]][2], ")", sep=""))+
    theme(axis.text.x = element_text(size = 14),
      text = element_text(size=20))

p1.Shp<-ggplot(Shape.lda.data, aes(LD1, LD2)) +
    geom_point(aes(color = Mode)) +
    theme_classic()+
    stat_ellipse(aes(color = Mode), type = "t", level = 0.75)+
    scale_color_brewer(palette="Dark2")+
    labs(color = "Shape",
         x = paste("LD1 (", LDShape.Coef[[4]][1], ")", sep=""),
         y = paste("LD2 (", LDShape.Coef[[4]][2], ")", sep=""))+
    theme(axis.text.x = element_text(size = 14),
      text = element_text(size=20))

p1.Mod<-ggplot(Mode.lda.data, aes(LD1, LD2)) +
    geom_point(aes(color = Mode)) +
    theme_classic()+
    stat_ellipse(aes(color = Mode), type = "t", level = 0.75)+
    scale_color_brewer(palette="Dark2")+
    labs(color = "Mode",
         x = paste("LD1 (", LDMode.Coef[[4]][1], ")", sep=""),
         y = paste("LD2 (", LDMode.Coef[[4]][2], ")", sep=""))+
    theme(axis.text.x = element_text(size = 14),
      text = element_text(size=20))

ggsave("LDACoefPlot_Habitat.eps", plot = p1.Hab, width = 15, height = 8, units = ("cm"))
ggsave("LDACoefPlot_Mode.eps", plot = p1.Mod, width = 15, height = 8, units = ("cm"))
ggsave("LDACoefPlot_Shape.eps", plot = p1.Shp, width = 15, height = 8, units = ("cm"))
```

# Morphology plot
```{r}
greys<-gray.colors(5, start = 0, end = 0.75)

morphPlot <- filter(vertMeasurementsAll, (Species == "Ap.flavidus" |
                                          Species == "Au.flavidus" |
                                          Species == "C.aggregata" |
                                          Species == "R.jordani"))

p1<- ggplot(data = morphPlot, aes(x = Position, y = CBL, color = Species))+
  geom_point()+
  geom_smooth(method = "lm", se = FALSE, formula = y ~ x + I(x^2))+
  xlab("Position") + ylab("Centra Length (BL)")+
  scale_color_manual(values = greys)+
  theme_classic()+
  labs(color = "Species")+
  theme(axis.text.x = element_text(size = 14),
      text = element_text(size=18),
      legend.position = "none")

p2<- ggplot(data = morphPlot, aes(x = Position, y = d, color = Species))+
  geom_point()+
  geom_smooth(method = "lm", se = FALSE, formula = y ~ x + I(x^2))+
  xlab("Position") + ylab("Foramen diameter (BL))")+
  scale_color_manual(values = greys)+
  theme_classic()+
  labs(color = "Species")+
  theme(axis.text.x = element_text(size = 14),
      text = element_text(size=18),
      legend.position = "none")

p3<- ggplot(data = morphPlot, aes(x = Position, y = alphaPost, color = Species))+
  geom_point()+
  geom_smooth(method = "lm", se = FALSE, formula = y ~ x + I(x^2))+
  xlab("Position") + ylab("Posterior cone angle (deg)")+
  scale_color_manual(values = greys)+
  theme_classic()+
  labs(color = "Species")+
  ylim(45,150)+
  theme(axis.text.x = element_text(size = 14),
      text = element_text(size=18))

plot_grid(p1, p2, p3, labels = "auto", ncol = 3, rel_widths = c(1,1,2))

ggsave("VariationInMorphology03.eps", width = 8.5, height = 2, units = "in")
```

# Read in Larger Fish Dataset for Tree Plot
```{r}
tree <- read.tree(file ="actinopt_12k_treePL.tre")

# removes species from dataset that are not in tree
my.data <- my.data[my.data$Species %in% tree$tip.label,] 

tree <-drop.tip(tree,tree$tip.label[-match(my.data$Species, tree$tip.label)]) 
# prunes tree to only include species in your data
```

# Clean and rotate data
```{r}
my.data.rotated <- filter(my.data, Pos != "20")
my.data.rotated <- filter(my.data.rotated, Pos != "30")
my.data.rotated <- filter(my.data.rotated, Pos != "40")
my.data.rotated <- filter(my.data.rotated, Pos != "50")

my.data.rotated <-
  my.data.rotated %>%
  select(c('Species', 'Habitat_FishBase', 'Pos', 'CBL', 'd', 'alphaPost', 'alphaAnt', 'Dpost', 'Dant', 'PercSquish',"predNmm"))

my.data.rotated <-
  my.data.rotated %>%
  group_by(Species) %>%
  mutate_at(vars(CBL, d, alphaPost, alphaAnt, Dpost, Dant, PercSquish, predNmm), ~ replace(.x, is.na(.x), mean(.x, na.rm=TRUE))) %>%
  ungroup()

my.data.rotated <-
my.data.rotated %>%
  pivot_wider(names_from = Pos, values_from = c('CBL', 'alphaPost', 'alphaAnt', 'd', 'Dpost','Dant', 'PercSquish',"predNmm"))
```

# Make Predictions based on Dana's tree Data
```{r}
treePredictions <- function(data){
  preproc.param <- data %>% preProcess(method = c("center", "scale"))
  transformed <- preproc.param %>% predict(data, method = "predictive")
  
  # Fit the model
  model <- lda(Habitat_FishBase~., data = transformed)
  
  testPredictions<- predict(model, transformed, method = "predictive")
  accuracy<- mean(na.omit(testPredictions$class==transformed$Habitat_FishBase))
  
  out <- list(model, testPredictions, accuracy)
}

treeLM.Linear <- treePredictions(my.data.rotated[c(2:26)])
treeLM.Perc <- treePredictions(my.data.rotated[c(2:30)])
treeLM.Pred <- treePredictions(my.data.rotated[c(2:26,31:34)])

# treeLM.Linear[[3]]
# treeLM.Perc[[3]]
# treeLM.Pred[[3]]

## Without Function = 0.7123
## With Perc Squish = 0.7945
## With predNmm = 0.7945205

habitatPredictions<- as.data.frame(treeLM.Perc[[2]]$class)
habitatKnown <- my.data.rotated$Habitat_FishBase
```

# Tree Plot
NOTE: Sometimes you get "Null" in the terminal when you run this chunk. If you do, just re-run one line at a time.
```{r}
pdf("HabTree.pdf", 15, 15)

x = as.matrix(habitatPredictions)
y = as.matrix(habitatKnown)
Y = cbind(x,y)

# cols<-setNames(RColorBrewer::brewer.pal(n=8,"Accent"),
#     unique(as.vector(sapply(as.factor(Y),levels))))
rownames(Y) <- (my.data.rotated$Species)
dotTree(tree,Y)

dev.off()
```

# Real Vert Bendint Plot
```{r}
realData <- filter(vertData, Type == "Real")
greys<-gray.colors(5, start = 0, end = 0.85)

f1 <- ggplot(realData, aes(x = PercSquish, y = Npermm))+
  geom_jitter(size = 2, aes(color = Species))+
  scale_color_manual(values = greys)+
  geom_smooth(method = "lm", se = FALSE)+
  theme_classic()+
  theme(text = element_text(size=10))+
  xlab("S:H ratio") + ylab("Force to break (N/mm)")

ggsave("ForceToBreak.eps", f1, width = 4, height = 2, units = "in")
```

